
### A Bigger, Better Data Set
Notes:

1. Use a Python script to get the current diamond price data, similar to the original diamonds data set.
2. Features of diamonds from all over the world.


```{r A Bigger, Better Data Set}
#install.packages('bitops')
#install.packages('RCurl')
library(ggplot2)
library(bitops)
library(RCurl)

load("BigDiamonds.rda")

str(diamondsbig)
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

Your task is to build five linear models like Solomon did for the diamonds data set only this time you'll use a sample of diamonds from the diamondsbig data set.

Be sure to make use of the same variables (logprice, carat, etc.) and model names (m1, m2, m3, m4, m5).

To get the diamondsbig data into RStudio on your machine, copy, paste, and run the code in the Instructor Notes. There's 598,024 diamonds in this data set!

Since the data set is so large, you are going to use a sample of the data set to compute the models. You can use the entire data set on your machine which will produce slightly different coefficients and statistics for the models.

```{r}
diamondsbig$logprice <- log(diamondsbig$price)
```

Setting price by diamonds price is less than 10000 dollars and certificate is G.I.A.

```{r Building a Model Using the Big Diamonds Data Set}
#Be sure you've loaded the library memisc and have m5 saved as an object in your workspace.
library(memisc)
m1 <- lm(logprice ~ I(carat^(1/3)), 
         data = diamondsbig[diamondsbig$price < 10000 &
                              diamondsbig$cert =="GIA",])
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits = 3)
```

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

We need to exponentiate the result of our model, since we took the log of price.

```{r}
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
exp(modelEstimate)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

The results yield an expected value for price, given the characteristics of our diamond and the upper and lower bounds of a 95% confidence interval. This is a linear model, predict is just multiplying each model coefficient by each value in our data. Maybe our source of the data is 


We see that the model predicts that the price for the diamond is 5040, while the actual price is 5601 dollars. The prediction is in our confidence interval.

The prediction interval here may be slightly conservative, as the model errors are heteroskedastic over carat (and hence price) even after our log and cube-root transformations. BlueNile has a better reputation than `http://www.diamondse.info/` So, while this model might give you a sense of wether your diamond is a ripoff against `http://www.diamondse.info/`, this shouldn't be regarded as the universal sourve of thruth about wether the price of a diamond is reasonable.

See the output of the following code.

```{r}
dat = data.frame(m4$model, m4$residuals)

with(dat, sd(m4.residuals))

with(subset(dat, carat > .9 & carat < 1.1), sd(m4.residuals))

dat$resid <- as.numeric(dat$m4.residuals)
ggplot(aes(y = resid, x = round(carat, 2)), data = dat) +
  geom_line(stat = "summary", fun.y = sd)
```

You can use a model like this to find out where you are overpaying. A diamond at [Tiffany is more expensive than at Costo](https://www.bloomberg.com/news/articles/2013-05-06/tiffany-vs-dot-costco-which-diamond-ring-is-better) but But Costco Sells Pricy Diamonds Too

[How to analyze your Facebook friends network with R](https://blog.revolutionanalytics.com/2013/11/how-to-analyze-you-facebook-friends-network-with-r.html)