
Lesson 7 Explore Many Variables
========================================================

Practice using multivariate techniques to look at more complicated relationships between multiple variables, and find out how to reshape your data.

## Price Histograms with Facets and Color

- Create a histogram of diamond prices.
- Facet the histogram by diamond color and use cut to color the histogram bars.
- Use a color palette of type 'qual' was used to color the histogram using `scale_fill_brewer(type = 'qual')`

```{r cars}
library(ggplot2)

data(diamonds)
#theme_set(theme_minimal(10))
summary(diamonds$price)
```

```{r Price Histograms with Facets and Color}
ggplot(aes(x = log(price)), data = diamonds) +
  geom_histogram(aes(color = cut)) +
  scale_fill_brewer(type = 'qual') +
  facet_wrap(~ color, ncol = 3) 
```


***

## Price vs. Table Colored by Cut

Create a scatterplot of diamond price vs.table and color the points by the cut of the diamond.

```{r Price vs. Table Colored by Cut}
ggplot(aes(x = table, y = price), data = diamonds) +
  geom_point(aes(color = cut)) +
  scale_fill_brewer(type = 'qual') +
  scale_x_continuous(breaks = seq(50, 80, 2)) +
  coord_cartesian(xlim = c(50, 80))
```


***

## Tipical Table Value

What is the typical table range for the majority of diamonds of ideal cut? 53-57


What is the typical table range for the majority of diamonds of premium cut? 58-62

***

## Price vs. Volume and Diamond Clarity

- Create a scatterplot of diamond price vs. volume (x * y * z) and color the points by the clarity of diamonds. 
- Use scale on the y-axis to take the log10 of price. 
- You should also omit the top 1% of diamond volumes from the plot.

```{r}
# create volume variable in the dataframe
diamonds <- transform(diamonds, volume = x * y * z)
ggplot(aes(x = volume, y = price), data = diamonds) +
  geom_point(aes(color = clarity)) +
  scale_color_brewer(type = 'div') +
  xlim(0, quantile(diamonds$volume, 0.99)) +
  scale_y_log10()
```

***

## Price/Carat Binned, Faceted & Colored

Create a scatter plot of the price/carat ratio of diamonds. The variable x should be assigned to cut. The points should be colored by diamond color, and the plot should be faceted by clarity.

```{r Price/Carat Binned, Faceted & Colored}
ggplot(aes(x = cut, y = price/carat), data = diamonds) +
  geom_point(aes(color = color)) +
  scale_color_brewer(type = 'div') +
  facet_wrap(~ clarity, ncol = 2) 
```


***