Explore Two Variables
========================================================

Practice using bivariate exploration techniques to explore relationships.

### Price versus x.

Create a scatterplot of `price` versus `x`.

```{r Price vs x Scatterplot}
library(ggplot2)
data(diamonds)
summary(diamonds$price)
summary(diamonds$x)
ggplot(aes(x = x, y = price), data = diamonds) +
  geom_point() 
```

From this scatterplot we can see that there are some outliers in the data set. Also, the relationship between `price` and `x` is exponential.

***
### Correlations

Let's look at the correlation coefficient to check the linear relationship between `x` and `price` variables.

```{r Correlations}
with(diamonds, cor.test(x, price))
with(diamonds, cor.test(y, price))
with(diamonds, cor.test(z, price))
```

## Correlation between `price` and `x`?
Response: 0.8844352

## Correlation between `price` and `y`?
Response: 0.8654209

## Correlation between `price` and `z`?
Response: 0.8612494 

```{r Correlation Coefficient}
ggplot(aes(x = x, y = price), data = diamonds) +
  coord_cartesian(xlim = c(3.5, quantile(diamonds$x, 0.90)), 
                  ylim = c(326, quantile(diamonds$price, 0.9))) +
  geom_point(alpha = 1/20, 
             position = position_jitter(h = 0),
             color = 'orange')  +
  geom_smooth(method = 'lm', color = 'red')
```


***

### Price versus Depth

```{r Price vs Depth}
ggplot(aes(x = depth, y = price), data = diamonds) +
  geom_point() 
```



***

### Adjustments - price vs depth 

```{r Price vs Depth Adjustments}
#summary(diamonds$price)
summary(diamonds$depth)
ggplot(aes(x = depth, y = price), data = diamonds) +
  geom_point(alpha = 1/10,
             position = position_jitter(h = 0),
             color = 'orange') +
  scale_x_continuous(breaks = seq(43, 79, 3))
```

***

### Typical Depth Range

Based on the scatterplot of Depth versus Price we can see tat most diamond are between 61 to 62.5 values of depth.

```{r Price vs Depth Range}
ggplot(aes(x = depth, y = price), data = diamonds) +
  geom_point(alpha = 1/20,
             position = position_jitter(h = 0),
             color = 'orange') +
  scale_x_continuous(breaks = seq(61, 62.5, 0.3)) +
  coord_cartesian(xlim = c(61, 62.5))
```

***

### Correlation Price and Depth

```{r Price and Depth Correlation}
cor.test(diamonds$depth, diamonds$price)
```

## What's the correlation of depth vs price?
Response: -0.01

Our intuition says no based on the plot of depth vs. plot and the correlation coefficient. This is because there isn't any linear relationship bewteen the two variables.

***

### Scatterplot of Price vs Carat

We will omit the top 1% of price and carat values.
```{r Price vs Carat}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point(alpha = 1/20,
             position = position_jitter(h = 0),
             color = 'orange') +
  scale_x_continuous(breaks = seq(0, 3, 0.1)) +
  coord_cartesian(xlim = c(0.15, quantile(diamonds$carat, 0.99)),
                  ylim = c(0, quantile(diamonds$price, 0.99)))
```

***

### Scatterplot of price vs. volume `(x * y * z)`

- This is a very rough approximation for a diamond's volume.
- Create a new variable for volume in the diamonds data frame.
- We won't make any adjustments to the plot just yet.

```{r Price vs. Volume}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(aes(x = volume, y = price), data = diamonds) +
  geom_point()
```

There are some extreme values in the data set. We can see that there are some volumes equal to zero. There is an expensive diamond with a volume near 4000 and a cheaper diamond with a volume near 900.

## Outliers

Let's see how many diamonds have a volume equal to zero.

```{r Zero Volume}
dim(subset(diamonds, volume == 0))[1]
```

We can see there are 20 data points with volume equal to 0.

***

### Correlations on Subsets

Let's see what's tha correlation of price and volume. I will exclude diamonds that have a volume of 0 or that are greater than or equal to 800.

```{r}
with(subset(diamonds, volume != 0 & volume <= 800),
     cor.test(volume, price))
```

***

### Adjustments Price vs Volume

- Subset the data to exclude diamonds with a volume  greater than or equal to 800. 
- Also, let's exclude diamonds with a volume of 0. 
- Adjust the transparency of the points and add a linear model to the plot. 
- Types of [smoothers](https://stats.idre.ucla.edu/r/faq/how-can-i-explore-different-smooths-in-ggplot2/) in `ggplot2`.

```{r Adjustments Price vs Volume}
ggplot(aes(x = volume, y = price), 
       data = subset(diamonds, volume != 0 & volume <= 800)) +
  geom_point(alpha = 1/10,
             position = position_jitter(h = 0),
             color = 'orange')  +
  ylim(0, 20000) +
  geom_smooth(method = 'lm', color = 'red', size = 1)
```

### Mean Price by Clarity

- Use the function `dplyr` package to create a new data frame containing info on diamonds by clarity.
- Name the data frame `diamondsByClarity`
- he data frame should contain the following variables in this order.
       (1) mean_price
       (2) median_price
       (3) min_price
       (4) max_price
       (5) n

```{r Mean Price by Clarity}
library(dplyr)
# group data frame by clarity
clarity_groups <- group_by(diamonds, clarity)
# summarize the clarity_groups in the new data frame
diamondsByClarity <- summarise(clarity_groups,
                               mean_price = mean(price),
                               median_price = median(price),
                               min_price = min(price),
                               max_price = max(price),
                               n = n())
# sort it by clarity
diamondsByClarity <- arrange(diamondsByClarity, clarity)
head(diamondsByClarity)
```

***

### Bar Charts of Mean Price

[Bar Charts vs. Histograms](https://flowingdata.com/2014/02/27/how-to-read-histograms-and-use-them-in-r/)

```{r Mean Price}
diamonds_mp_by_clarity <- summarise(clarity_groups, mean_price = mean(price))
diamonds_mp_by_clarity <- arrange(diamonds_mp_by_clarity, mean_price)
diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
diamonds_mp_by_color <- arrange(diamonds_mp_by_color, mean_price)
head(diamonds_mp_by_clarity)
head(diamonds_mp_by_color)
```

Let's create two Box Plots with the mean price by `clarity` and `color`.


```{r Bar Charts of Mean Price}
p1 <- barplot(diamonds_mp_by_clarity$mean_price, 
        names.arg = diamonds_mp_by_clarity$clarity,
        main = 'Diamond Price by Clarity')

p2 <- barplot(diamonds_mp_by_color$mean_price, 
        names.arg = diamonds_mp_by_color$color,
        main = 'Diamond Price by Color')

#library(gridExtra)

#grid.arrange(p1,p2, ncol=1)


```

***
### Trends in Mean Price

In the Diamond Price by Clarity box chart we can see that the average price for diamonds is cheaper for a VVS1 clarity diamond and most expensive for a SI2 clarity diamond. We can see that the average diamond price is similar, just under 4000$, for a clarity of VS1, I1, VS2 or SI1.
In the Diamond Price by Colors box chart we can see that colors for diamonds are in letters ranging from E to J with least expensive for E, at around 3000 dollars and most expensive for J, more than 5000 dollars.

- We think something odd is going here. These trends seem to go against our intuition.
- Mean price tends to decrease as clarity improves. The same can be said for color.
- We encourage you to look into the mean price across cut.